---
title: "Research project"
author: "470535297"
output:
  pdf_document: default
  word_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Getting the Data files:
  First of all we need to read the given three data files which is in CSV format. We want to bring the files in the R data frame in order to read and study the domestic violance happening in the NSW Local LGA. The NSW LGA file is having the Data of domestic violance monthly basis from january 99 to december 15. This code is given bellow:
  
\vspace{3mm}

```{r}
# code for upload the initial data files
df_dv_nsw<- read.csv("C:/Users/golam/Desktop/assesment-data/data/DV_NSW_by_LGA.csv")
```

\vspace{6mm}
  The NSW Lga file is having the data of various NSW LGA in 2011 data where we can see the data of various catogery is exists. the code is given bellow to read the file:

\vspace{3mm}
```{r}
df_nsw_lga<- read.csv("C:/Users/golam/Desktop/assesment-data/data/NSW_LGA.csv")
```
\vspace{6mm}
Finally we have the data file For the labels where the description is given about the data pack, such as B1 is the total number  of male. the code is given bellow:

\vspace{3mm}

```{r}

df_lebel<- read.csv("C:/Users/golam/Desktop/assesment-data/data/labels.csv")
```

\vspace{3mm}

Read to analyse the data geting from the NSW Recorded Crime Statistics January to December 2017.

```{r}
library(readxl)
otherdata<- read_excel("C:/Users/golam/Desktop/assesment-data/data/otherdata.xls",
                       sheet="Location")

#df[!is.na(df$B), ]
otherdata[!is.na(otherdata$'Incident Local Government Area'), ]
otherdata[!is.na(otherdata$'Number of incidents'), ] 

```

\vspace{3mm}

# Exploring the data files:

## Attribute Information

### Here is the attribute information for our three data file which is given from the censes data of NSW government data set:

\vspace{3mm}

####DV_NSW_by_LGA:
The  given DV_NSW_by_LGA file is containing the data of 140 local government areas data of domestic violance as a time series data starting from january 99 till december 2015. Variables are noted bellow:

    1) LGA: Name of all the 140 LGA. 
    2) All other colums containing the number of incedent happen in every month.

\vspace{3mm}

####NSW_LGA:
This file contain the data of all incedents including the area and catagorical data. Variables are noted bellow:

    1) region_id: ID of the corresponding regional area. 
    2) label:This field contain the name of the Regeon.
    3) year: year of the incedent happened.
    4) area: size of the area.
    5) rest of all other field contains the data of number of people 
       based on the catogory of the labels.
    
\vspace{3mm}

####label:
This file is mainly important for the description of the labels. 

\vspace{3mm}

# Cleaning the data
Next we have to clean this data. This data is actually almost already cleaned for, But here are some things need to consider doing for other data sets:

###Check for NA values
Let's see if there any NA values:

```{r}
# checking is there any NULL values in the initial data files
any(is.na(df_dv_nsw))
```

```{r}
any(is.na(df_lebel))
```

```{r}
any(is.na(df_nsw_lga))
```

\vspace{3mm}
we have seen that all three files have no null values or the nill. However it is also important to see the catagorical feature of the data file before further analysis the data and also having a look the visualization of the data. 


\vspace{6mm}
Now lets see the head of three data files. bellow is the time series data of the NSW lga files.
\vspace{3mm}



\vspace{6mm}
Bellow is the head of the label file.
\vspace{3mm}

```{r}
head(df_lebel)
```

\vspace{6mm}

#Exploratory Data Analysis

In order to get a total picture of domestic violance happening in nsw local areas based on the data been supplied, we have to do certain type of operation within the data files and have to  combine some field for the exploratory data analysis. 

\vspace{3mm}

Since the data given in the NSW_LGA file is only consedering the data of 2011 so in the CSV file of DV_NSW ONLY 2011 data is added in a sepereate colum, aiming to analyse this data with the NSW_Lga file. the code for combining the fields is given bellow:


### Process for combining the data files.
Since the data in the domestic violance file is consist as a monthly data, this has been converted to yearly data for the better iteration.

```{r}

# Combining the data of 2011 in one variable for future iteration


df_dv_nsw$all_month_2011 <- df_dv_nsw$Jan.11 + df_dv_nsw$Feb.11 + df_dv_nsw$Mar.11 +  
  df_dv_nsw$Apr.11 + df_dv_nsw$May.11 + df_dv_nsw$Jun.11 + df_dv_nsw$Jul.11 + 
  df_dv_nsw$Aug.11+  df_dv_nsw$Sep.11 + df_dv_nsw$Oct.11 + 
  df_dv_nsw$Nov.11 + df_dv_nsw$Dec.11



# Combining the data of 99 in one variable for future iteration

df_dv_nsw$all_month_99 <- df_dv_nsw$Jan.99 + df_dv_nsw$Feb.99 + df_dv_nsw$Mar.99 +  
  df_dv_nsw$Apr.99 +  df_dv_nsw$May.99 + df_dv_nsw$Jun.99 + df_dv_nsw$Jul.99 + 
  df_dv_nsw$Aug.99+  df_dv_nsw$Sep.99 + df_dv_nsw$Oct.99 +  
  df_dv_nsw$Nov.99 + df_dv_nsw$Dec.99

# Combining the data of 00 in one variable for future iteration

df_dv_nsw$all_month_00 <- df_dv_nsw$Jan.00 + df_dv_nsw$Feb.00 + df_dv_nsw$Mar.00 +  
  df_dv_nsw$Apr.00 +  df_dv_nsw$May.00 + df_dv_nsw$Jun.00 + df_dv_nsw$Jul.00 + 
  df_dv_nsw$Aug.00+  df_dv_nsw$Sep.00 + df_dv_nsw$Oct.00 +  
  df_dv_nsw$Nov.00 + df_dv_nsw$Dec.00

# Combining the data of 01 in one variable for future iteration

df_dv_nsw$all_month_01 <- df_dv_nsw$Jan.01 + df_dv_nsw$Feb.01 + df_dv_nsw$Mar.01 +  
  df_dv_nsw$Apr.01 +  df_dv_nsw$May.01 + df_dv_nsw$Jun.01 + df_dv_nsw$Jul.01 + 
  df_dv_nsw$Aug.01+  df_dv_nsw$Sep.01 + df_dv_nsw$Oct.01 +  
  df_dv_nsw$Nov.01 + df_dv_nsw$Dec.01

# Combining the data of 02 in one variable for future iteration

df_dv_nsw$all_month_02 <- df_dv_nsw$Jan.02 + df_dv_nsw$Feb.02 + df_dv_nsw$Mar.02 +  
  df_dv_nsw$Apr.02 +  df_dv_nsw$May.02 + df_dv_nsw$Jun.02 + df_dv_nsw$Jul.02 + 
  df_dv_nsw$Aug.02+  df_dv_nsw$Sep.02 + df_dv_nsw$Oct.02 +  
  df_dv_nsw$Nov.02 + df_dv_nsw$Dec.02

# Combining the data of 03 in one variable for future iteration

df_dv_nsw$all_month_03 <- df_dv_nsw$Jan.03 + df_dv_nsw$Feb.03 + df_dv_nsw$Mar.03 +  
  df_dv_nsw$Apr.03 +  df_dv_nsw$May.03 + df_dv_nsw$Jun.03 + df_dv_nsw$Jul.03 + 
  df_dv_nsw$Aug.03+  df_dv_nsw$Sep.03 + df_dv_nsw$Oct.03 +  
  df_dv_nsw$Nov.03 + df_dv_nsw$Dec.03

# Combining the data of 04 in one variable for future iteration

df_dv_nsw$all_month_04 <- df_dv_nsw$Jan.04 + df_dv_nsw$Feb.04 + df_dv_nsw$Mar.04 +  
  df_dv_nsw$Apr.04 +  df_dv_nsw$May.04 + df_dv_nsw$Jun.04 + df_dv_nsw$Jul.04 + 
  df_dv_nsw$Aug.04+  df_dv_nsw$Sep.04 + df_dv_nsw$Oct.04 +  
  df_dv_nsw$Nov.04 + df_dv_nsw$Dec.04

# Combining the data of 05 in one variable for future iteration

df_dv_nsw$all_month_05 <- df_dv_nsw$Jan.05 + df_dv_nsw$Feb.05 + df_dv_nsw$Mar.05 +  
  df_dv_nsw$Apr.05 +  df_dv_nsw$May.05 + df_dv_nsw$Jun.05 + df_dv_nsw$Jul.05 + 
  df_dv_nsw$Aug.05+  df_dv_nsw$Sep.05 + df_dv_nsw$Oct.05 +  
  df_dv_nsw$Nov.05 + df_dv_nsw$Dec.05

# Combining the data of 06 in one variable for future iteration

df_dv_nsw$all_month_06 <- df_dv_nsw$Jan.06 + df_dv_nsw$Feb.06 + df_dv_nsw$Mar.06 +  
  df_dv_nsw$Apr.06 +  df_dv_nsw$May.06 + df_dv_nsw$Jun.06 + df_dv_nsw$Jul.06 + 
  df_dv_nsw$Aug.06+  df_dv_nsw$Sep.06 + df_dv_nsw$Oct.06 +  
  df_dv_nsw$Nov.06 + df_dv_nsw$Dec.06

# Combining the data of 07 in one variable for future iteration

df_dv_nsw$all_month_07 <- df_dv_nsw$Jan.07 + df_dv_nsw$Feb.07 + df_dv_nsw$Mar.07 +  
  df_dv_nsw$Apr.07 +  df_dv_nsw$May.07 + df_dv_nsw$Jun.07 + df_dv_nsw$Jul.07 + 
  df_dv_nsw$Aug.07+  df_dv_nsw$Sep.07 + df_dv_nsw$Oct.07 +  
  df_dv_nsw$Nov.07 + df_dv_nsw$Dec.07

# Combining the data of 08 in one variable for future iteration

df_dv_nsw$all_month_08 <- df_dv_nsw$Jan.08 + df_dv_nsw$Feb.08 + df_dv_nsw$Mar.08 +  
  df_dv_nsw$Apr.08 +  df_dv_nsw$May.08 + df_dv_nsw$Jun.08 + df_dv_nsw$Jul.08 + 
  df_dv_nsw$Aug.08+  df_dv_nsw$Sep.08 + df_dv_nsw$Oct.08 +  
  df_dv_nsw$Nov.08 + df_dv_nsw$Dec.08

# Combining the data of 09 in one variable for future iteration

df_dv_nsw$all_month_09 <- df_dv_nsw$Jan.09 + df_dv_nsw$Feb.09 + df_dv_nsw$Mar.09 +  
  df_dv_nsw$Apr.09 +  df_dv_nsw$May.09 + df_dv_nsw$Jun.09 + df_dv_nsw$Jul.09 + 
  df_dv_nsw$Aug.09+  df_dv_nsw$Sep.09 + df_dv_nsw$Oct.09 +  
  df_dv_nsw$Nov.09 + df_dv_nsw$Dec.09

# Combining the data of 10 in one variable for future iteration

df_dv_nsw$all_month_10 <- df_dv_nsw$Jan.10 + df_dv_nsw$Feb.10 + df_dv_nsw$Mar.10 +  
  df_dv_nsw$Apr.10 +  df_dv_nsw$May.10 + df_dv_nsw$Jun.10 + df_dv_nsw$Jul.10 + 
  df_dv_nsw$Aug.10+  df_dv_nsw$Sep.10 + df_dv_nsw$Oct.10 +  
  df_dv_nsw$Nov.10 + df_dv_nsw$Dec.10

# Combining the data of 12 in one variable for future iteration

df_dv_nsw$all_month_12 <- df_dv_nsw$Jan.12 + df_dv_nsw$Feb.12 + df_dv_nsw$Mar.12 +  
  df_dv_nsw$Apr.12 +  df_dv_nsw$May.12 + df_dv_nsw$Jun.12 + df_dv_nsw$Jul.12 + 
  df_dv_nsw$Aug.12 +  df_dv_nsw$Sep.12 + df_dv_nsw$Oct.12 +  
  df_dv_nsw$Nov.12 + df_dv_nsw$Dec.12

# Combining the data of 13 in one variable for future iteration

df_dv_nsw$all_month_13 <- df_dv_nsw$Jan.13 + df_dv_nsw$Feb.13 + df_dv_nsw$Mar.13 +  
  df_dv_nsw$Apr.13 +  df_dv_nsw$May.13 + df_dv_nsw$Jun.13 + df_dv_nsw$Jul.13 + 
  df_dv_nsw$Aug.13 +  df_dv_nsw$Sep.13 + df_dv_nsw$Oct.13 +  
  df_dv_nsw$Nov.13 + df_dv_nsw$Dec.13

# Combining the data of 14 in one variable for future iteration

df_dv_nsw$all_month_14 <- df_dv_nsw$Jan.14 + df_dv_nsw$Feb.14 + df_dv_nsw$Mar.14 +  
  df_dv_nsw$Apr.14 +  df_dv_nsw$May.14 + df_dv_nsw$Jun.14 + df_dv_nsw$Jul.14 + 
  df_dv_nsw$Aug.14 +  df_dv_nsw$Sep.14 + df_dv_nsw$Oct.14 +  
  df_dv_nsw$Nov.14 + df_dv_nsw$Dec.14

# Combining the data of 15 in one variable for future iteration

df_dv_nsw$all_month_15 <- df_dv_nsw$Jan.15 + df_dv_nsw$Feb.15 + df_dv_nsw$Mar.15 +  
  df_dv_nsw$Apr.15 +  df_dv_nsw$May.15 + df_dv_nsw$Jun.15 + df_dv_nsw$Jul.15 + 
  df_dv_nsw$Aug.15 +  df_dv_nsw$Sep.15 + df_dv_nsw$Oct.15 +  
  df_dv_nsw$Nov.15 + df_dv_nsw$Dec.15

```

\vspace{3mm}
So from the above code it is targeting that the 2011 data is been examined along with the corresponding Area codes. 


\vspace{3mm}

Bellow is the code for selecting new colums.

```{r}
#code to keep the two variables in the new data file df_dv_nsw1
df_dv_nsw1 <- subset(df_dv_nsw, select = c(1, 206, 207, 208, 209, 210, 211, 212, 
                                           213, 214, 215, 216, 217, 218, 
                                           219, 220, 221, 222))

```
\vspace{3mm}


Now new file is become the df_dv_nsw1, which is being usable in future with the combination of the Nsw_lga file. Bellow is the code for LGA wise domestic violance data only for the year of 2011. Nice thing is this data file only contain 2 variables.

Bellow is the head of data which is bacically the total summary of domestic violance year to year.


\vspace{3mm}
```{r}
head(df_dv_nsw1)

```

\vspace{5mm}

### Summerising the yearly data to see the graph representation.

```{r}
dv99<-sum(df_dv_nsw1$all_month_99)
dv00<-sum(df_dv_nsw1$all_month_00)
dv01<-sum(df_dv_nsw1$all_month_01)
dv02<-sum(df_dv_nsw1$all_month_02)
dv03<-sum(df_dv_nsw1$all_month_03)
dv04<-sum(df_dv_nsw1$all_month_04)
dv05<-sum(df_dv_nsw1$all_month_05)
dv06<-sum(df_dv_nsw1$all_month_06)
dv07<-sum(df_dv_nsw1$all_month_07)
dv08<-sum(df_dv_nsw1$all_month_08)
dv09<-sum(df_dv_nsw1$all_month_09)
dv10<-sum(df_dv_nsw1$all_month_10)
dv11<-sum(df_dv_nsw1$all_month_2011)
dv12<-sum(df_dv_nsw1$all_month_12)
dv13<-sum(df_dv_nsw1$all_month_13)
dv14<-sum(df_dv_nsw1$all_month_14)
dv15<-sum(df_dv_nsw1$all_month_15)

dv=c(dv99,dv00, dv01, dv02, dv03, dv04, dv05, dv06, dv07, dv08, 
     dv09, dv10, dv11, dv12, dv13, dv14, dv15)
year=c(1999,2000,2001,2002,2003,2004,2005,2006,2007,
       2008,2009,2010,2011,2012,2013,2014,2015)

setNames(dv, year)
#as.table(setNames(dv, year))
data.frame(setNames(dv, year))


barplot(setNames(dv,year),xlab='Year from 99 to 15', ylab='Number of domestic violance', 
        main='Yearly domestic violance in nsw', col='lightblue')


```

\vspace{5mm}

From the above graph it has been seen that the trand of domestic violance in nsw is going up which is very significant to mention, and is so alerming. however since from 1999 the popolation in NSW is increasing, however we have to see and examine the corelation with the increasing population density in NSW as well. 

\vspace{5mm}

### Bellow is the visualization of the data file which we called: other data.

```{r}

#head(otherdata)
library(ggplot2)
names(otherdata)[names(otherdata) == "Incident Local Government Area"] <- "Area"
names(otherdata)[names(otherdata) == "Number of incidents"] <- "Incidents"
names(otherdata)[names(otherdata) == "Rate per 100,000"] <- "Rate"
#head(otherdata)

otherdata$Rank <- as.numeric(otherdata$Rank)

attach(otherdata)
newdata <- otherdata[ which(otherdata$Rank<11),]
#head(newdata)

p<-ggplot(data=newdata, aes(x=Area,y=Incidents)) +
   geom_bar(stat="identity", fill="steelblue")
 
p
  

```

Fig is showing top ten domestic violance rating area in NSW from the data 'NSW Recorded Crime Statistics January to December 2017.'


\vspace{5mm}
## Now lets visualise some of the data yearly basis rate of domestic violance area wise. 

```{r}
#Scatter plot of domestic violance in 2011

library(ggplot2)

plot_dv_2011 <- ggplot(data = df_dv_nsw1, mapping = aes(x = LGA , y = all_month_2011)) + 
  geom_point() +
  labs(title="ScatterPlot: DV 2011",x="LGA",y="count of DV")
plot_dv_2011


plot(all_month_2011~all_month_15, data=df_dv_nsw1)



```


\vspace{5mm}

NSW_LGA containing total 7946 variables, which not all them are required to analyse the domestic violance in nsw. Considering the examining year bellow is the variables, been selected to analyse. Bellow is the nine variables been selected for further analysis:

\vspace{3mm}

  - **region_id** ID of the region
  - **label**	Name of the area
  - **year** 2011
  - **area_sqkm** Area of that Region in squire kilometer
  - **B1** Total_Persons_Males
  - **B2**	Total_Persons_Females
  - **B3** Total_Persons_Persons
  - **B15** Age_groups_20_24_years_Persons
  - **B18** Age_groups_25_34_years_Persons
  
\vspace{5mm}

Bellow is the code for new selected data colums, where initially the variables been select is only 9. This new data file and dv_nsw1 file will be combined to get the final workable file to explain the domestic violance in nsw. This is noticable that the LGA variable field in this file has converted to the same shape as the DV_NSW, So that further merging is possible is possible in between the CSV files. The code and final structure of the data file is given bellow:

\vspace{5mm}

```{r}
# keeping some selected variables which i have the interest, rather 
#than keeping 7942 variables
df_nsw_lga1 <- subset(df_nsw_lga, select = c(1,2,3,4,5,6,7,19,22))
#setnames(df_nsw_lga1, "label", "LGA")
colnames(df_nsw_lga1)[which(names(df_nsw_lga1) == "label")] <- "LGA"
df_nsw_lga1$LGA <- gsub('.{4}$', '',df_nsw_lga1$LGA )
head(df_nsw_lga1)
```

\vspace{5mm}

Bsed on the LGA varible, the data file df_dv_nsw1 and df_dv_lga1 is been combined to the final workable data file in the NSW LGA area, to analyse and explain the domestic violance situation. The domestic violance is been considered is based on the data of the year 2011. The code and final data file head is given bellow:

\vspace{5mm}

library sqldf has been used in here, to select the matcheced colums from both of the csv file df_nsw_lga1 and df_dv_nsw1, whicle the matches colum is LGA from both of the file. the result has been stored in the new csv file called data.csv, which has been used in the rest of the analysis of the domestic violance probel. since two colums remain again the same name we need to drom one colum name "LGA". the final gigure of the data file is mentioned bellow.


```{r}
#Combine new two data file into one using the SQL library and statement, and put them in the final 
#datafile which have to be obsetrve later.

df1<-df_nsw_lga1
df2<-df_dv_nsw1


library(sqldf )
data<-sqldf("SELECT *
      FROM df1, df2
      WHERE df1.LGA==df2.LGA")

data <- subset(data, select = c(1,2,3,4,5,6,7,8,9,11))

head(data)

```


\vspace{5mm}
This new workable data file having the name of some variables, which need to change the name for better understand once observe the structure of the data file. 

\vspace{5mm}

```{r}
# changing the name of some variables and give the sensable name
names(data)[names(data) == "B1"] <- "maleTotal"
names(data)[names(data) == "B2"] <- "femaleTotal"
names(data)[names(data) == "B3"] <- "populationTotal"
names(data)[names(data) == "B15"] <- "Twenty_twentyfour_Total"
names(data)[names(data) == "B18"] <- "twentyfive_thirtyfive_Total"
```

\vspace{5mm}
```{r}
head(data)
```

\vspace{5mm}

# Further explanation from this point

\vspace{5mm}

```{r}
# checking is there any NULL value in our new data file
any(is.na(data))
```

\vspace{5mm}
```{r}

# The summary of data of the new data file
summary(data)
```
\vspace{5mm}

## Finding the co-orelation in between the numaric colum in this table is given bellow.

Finding the co orelation between the data is one of the important task in statistical learning. bellow is the explanation of the corelation in the data field of this table:

Correlation plots are a great way of exploring data and seeing if there are any interaction terms. 

```{r}


data$year <- as.character(data$year)
# Grab only numeric columns
num.cols <- sapply(data, is.numeric)

# Filter to numeric columns for correlation
cor.data <- cor(data[,num.cols])

cor.data
```

\vspace{5mm}

```{r}
library(corrplot)
library(corrgram)

#Visualization the relation within the variables.

corrgram(data,order=TRUE, lower.panel=panel.shade,
  upper.panel=panel.pie, text.panel=panel.txt)


plot(all_month_2011~area_sqkm, data=data)
plot(all_month_2011~maleTotal, data=data)
plot(all_month_2011~populationTotal, data=data)
plot(all_month_2011~femaleTotal, data=data)

```


The pie in the figure shows that the blue area is perfectly co orelated where as the other color is mean that not perfectly corelated. 

\vspace{5mm}

### Since we're going to eventually try to see the domestic violance in the year of 2011, lets see the histogram of it:

\vspace{5mm}

```{r}
library(ggplot2)
ggplot(data,aes(x=all_month_2011)) + geom_histogram(bins=20,alpha=0.5,fill='blue') + theme_minimal()

```






\vspace{5mm}

# Building model
```{r}


model <- lm(all_month_2011 ~area_sqkm+populationTotal+maleTotal+femaleTotal+
              Twenty_twentyfour_Total+twentyfive_thirtyfive_Total ,data)


```
\vspace{6mm}

## Model summary

```{r}
summary(model)
```

##Model description and analysis/result:
  #### R-Squire shows the model in a good fitting with the variable we are predicting. 

  ### From the probability of co efficient we can see, in terms of popolation and the male population the model is denoted very significance. which mean that the number of domesticviolance in NSW is  really depends on the populatin or on the male, population.  The standered error of the model is significantly low, which denotes that the model predecting a good result,which is very remarkable. the domestic violance is than may be related to the other factor like the income of the house hold or may be the educational levels of the indivisuals. where as we dont have that data available to analyse. 


\vspace{5mm}

## Model visualization

```{r}
# Grab residuals
res <- residuals(model)

# Convert to DataFrame for gglpot
res <- as.data.frame(res)

head(res)
```


##Using ggplot

```{r}
# Histogram of residuals

library(ggplot2)
ggplot(res,aes(res)) +  geom_histogram(fill='blue',alpha=0.5)

```


\vspace{5mm}
The above distribution of Residual shows that the model we made was preety good as the distribution is almost 
Gausian or the normal distribution. 


\vspace{5mm}

## plot model

```{r}

plot(model)

```






